---
title: "Simulations Pre-Class Project"

output:
  html_document


---

<style type="text/css">
.table {

    width: 80%;
    margin-left:10%; 
    margin-right:10%;
}
</style>
```{r,setup, echo=FALSE, cache=TRUE}
## numbers >= 10^5 will be denoted in scientific notation,
## and rounded to 2 digits
options(scipen = 3, digits = 3)
```




#Project Goals:


With this project we will simulate a famoues probability problem. This will not require knowledge of probability or statistics but only the logic to follow the steps in order to simulate this problem. This is one way to solve problems by using the computer. 

 1. **Gambler's Ruin**: Suppose you have a bankroll of $1000 and make bets of $100 on a fair game. By simulating the outcome directly for at most 5000 iterations of the game (or hands), estimate:
    a. the probability that you have "busted" (lost all your money) by the time you have placed your one hundredth bet. 
```{r}
# this function will return when the player busted
whenBust=function()
{
  iteration=1
  fraction=100/1000
  bet=100
  bankroll=1000
  while (bankroll>0  & iteration<5000) # since that player play at most 5000 hands
  {
    winorlose=rbinom(1,1,0.5)
    if (winorlose==1)
    {
      bankroll=bankroll+bet
      bet=bankroll*fraction
    }else
    {
      bankroll=bankroll-bet
    }
    iteration=iteration+1
    
    }  
  
  return (iteration)
}

simulations=rep(NA,5000)
for (i in 1:5000)
{
  simulations[i]=whenBust()
}

length(simulations[simulations<=100])/5000

```

   b. the probability that you have busted by the time you have placed your five hundredth bet by simulating the outcome directly.
```{r}
for (i in 1:5000)
{
  simulations[i]=whenBust()
}
length(simulations[simulations<=500])/5000

```

    
  c. the mean time you go bust, given that you go bust within the first 5000 hands.
```{r}
for (i in 1:5000)
{
  simulations[i]=whenBust()
}
mean(simulations)
```


  d. the mean and variance of your bankroll after 100 hands (including busts).
```{r}
# the input（x）of this fucntion is an iteration time (when the game stop), and will return Bankroll after that iteration time
simbankroll=function(x)
{
  iteration=1
  fraction=100/1000
  bet=100
  bankroll=1000
  while (iteration<=x & bankroll>0)
  {
     winorlose=rbinom(1,1,0.5)
    if (winorlose==1)
    {
      bankroll=bankroll+bet
      bet=bankroll*fraction
    }else
    {
      bankroll=bankroll-bet
    }
    iteration=iteration+1
    } 
  return (bankroll)
}


sim_bankroll=rep(NA,5000)
for (i in 1:5000)
{
  sim_bankroll[i]=simbankroll(100)
}

mean(sim_bankroll)
var(sim_bankroll)
```

  e. the mean and variance of your bankroll after 500 hands (including busts).
```{r}
sim_bankroll2=rep(NA,5000)

for (i in 1:5000)
{
  sim_bankroll2[i]=simbankroll(500)
}
mean(sim_bankroll2)
var(sim_bankroll2)
```

 
Note: you *must* stop playing if your player has gone bust. How will you handle this in the `for` loop?
