---
title: "pre-class-03.Rmd"
author: "Chengqian Sun"
date: "2018/9/25"
output: html_document
---

# pre-class

Make sure you commit this often with meaningful messages. 


### Getting Started


We will work with the dataset called [gapminder](https://github.com/jennybc/gapminder), this is a cleaned up version from [Gapminder Data](http://www.gapminder.org/data/). Gapminder contains a lot of great data on all of the nations of the world. We first need to install the gapminder package in R. 






```{r}
library(dplyr)
library(gapminder)
gapminder
```




### Pre-Class Problems

Use **dplyr** functions to address the following questions:

1. How many unique countries are represented per continent?
```{r}
#first group by continent then find the number of distinct countries in each continent 
gapminder %>%
group_by(continent) %>%
summarize(NumofCountry = n_distinct(country))

```


2. Which European nation had the lowest GDP per capita in 1997? 

```{r}
# first find the countries in EU and the year is at 1997, then arrange them.

gapminder %>%
 filter(continent=="Europe"& year=="1997")%>%
 arrange(gdpPercap)



```

The country of first row  Albania, thus the Albania has lowest gdp in EU at 1997

3. According to the data available, what was the average life expectancy across each continent in the 1980s?
```{r}
#first find the rows in 80s, then group them by continent. Then calculate the mean of life expectancy of each continent.

gapminder %>%
 filter(year>=1980 & year<=1989)%>%
 group_by(continent)%>%
 summarize(avg_lfe=mean(lifeExp))
```


4. What 5 countries have the highest total GDP over all years combined?

```{r}
# group rows by contries, then calcluate the total gdp for each country. then arrange these countries, find the highest 5.

gapminder%>%
  group_by(country)%>%
  summarize(tot_gdp=sum(gdpPercap))%>%
  arrange(desc(tot_gdp))

```

We can see that the first 5 countries are Kuwait, Switzerland, Norway,United States and Canada.

5. What countries and years had life expectancies of _at least_ 80 years? _N.b. only output the columns of interest: country, life expectancy and year (in that order)._

```{r}
gapminder%>%
  filter(lifeExp>=80)%>%
  select(country,lifeExp,year)


```

6. What 10 countries have the strongest correlation (in either direction) between life expectancy and per capita GDP?
```{r}
gapminder%>%
  group_by(country)%>%
  summarize(cor=abs(cor(lifeExp,gdpPercap)))%>%
  arrange(desc(cor))

```



7. Which combinations of continent (besides Asia) and year have the highest average population across all countries? _N.b. your output should include all results sorted by highest average population_. With what you already know, this one may stump you. See [this Q&A](http://stackoverflow.com/q/27207963/654296) for how to `ungroup` before `arrange`ing. This also [behaves differently in more recent versions of dplyr](https://github.com/hadley/dplyr/releases/tag/v0.5.0).


```{r}

gapminder%>%
  filter(continent!="Asia")%>%
  group_by(continent,year)%>%
  summarize(avg_pop=mean(pop))%>%
  arrange(desc(avg_pop))

```
We can see that the combinaition of having the highest average population across all countries is Americas and 1997

8. Which three countries have had the most consistent population estimates (i.e. lowest standard deviation) across the years of available data? 

```{r}
gapminder%>%
  group_by(country)%>%
  summarize(sd=sd(pop))%>%
  arrange(sd)
  
  

```
We can see that the three countries with smallest standard deviation are:
Sao Tome and Principe, Iceland, Montenegro


9. Subset **gm** to only include observations from 1992 and store the results as **gm1992**. What kind of object is this?

```{r}
gm1992=filter(gapminder,year==1992)

```
The gm1992 is a subset of gapminder, thus they are same type of object. If gapminder is a tibble, then gm1992 is a tibbe

10. Which observations indicate that the population of a country has *decreased* from the previous year **and** the life expectancy has *increased* from the previous year? See [the vignette on window functions](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html).


```{r}


gapminder%>%
  filter(pop<lag(pop) & lifeExp>lag(lifeExp))


```



