# pre-class


Make sure you commit this often with meaningful messages. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

```
f1 <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}
```
The input parameters of this function are two strings, and it return a
boolean variable.
f1 (a,b) is the function to judge that whether b is a prefix of a.
Which means to judge if one string is the beginning of the other string.
I would like to call f1 as "ifPrefix"

```
f2 <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}
```
f2(x), return an vector that removes the last object of x. However, if there
are less or equal to 1 object in x. f2(x) will return NULL
I would like to call this function as "RemoveLast"
```
f3 <- function(x, y) {
  rep(y, length.out = length(x))
}
```

f3(x,y) will return a vector that has same length of x, and all obejects in that vector 
are objects of y. If x is shorter than y, it will return the first n=length(x) objects 
of y. If x is longer than y, the returned vector will recycle the y.
I would like to call it "ReplicateY"


2. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?
```{r}
?rnorm()
?mvnorm()
```

Based on the description of these two functions.
`rnorm(n,mean,sd)`generates n random sampels from normal distribution with mean=mean, and standard
devaition=sd. 
`mvnorm(n,mu,Sigma)` generates n samples from multivariate normal distribution. Where mu is a vector giving the means of the variables, and Sigma is a positive-definite symmetric matrix specifying the covariance matrix of the variables.
We can see that these two functions generates random samples based on different distribution.
and 'sd' in `rnorm()` is a vector of sandarad devations. However, 'Sigma' in `mvrnorm()` is
positive-definite symmetric matrix specifying the covariance matrix of the variables, which is
not a standarad devation. 
To make this two functions more consistent. When we use `mvnorm()`, we can let 'mu' equals to
'mean' in rnorm(), and 'Sigma' equals to sd^2, where sd is standard devation in `rnorm`

```{r}
rn=rnorm(1000,5,4)
mn=mvrnorm(1000,5,16)
qqplot(rn,mn)
#based on qqnorm, the samples generated from these two function are simmilar.

```

3. Use `lapply()` and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset. 

```{r}
lapply(mtcars, function(x){mean(x)/sd(x)})
```

4. Use vapply() to:
    a. Compute the standard deviation of every column in a numeric data frame.
```{r}
#4(a)
vapply(mtcars, function(x) {sd(x)}, numeric(1))
```
     b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)
```{r}
#4(b)
# first build a dataframe that contains non-numeric varaible
string=rep("hello",nrow(mtcars))
newdata=cbind(mtcars,string)
newdata
vapply(newdata[vapply(newdata, is.numeric, logical(1))], sd, numeric(1))
```

